An assessment of photometric redshift PDF performance in the context of LSST

### 2018-11-07 paper/tex/methods.tex
address reviewer comments in method section

\section{Methods}
\label{sec:pzcodes}

%(Sam Schmidt, Rongpu Zhou, Ibrahim Almosallam, Eric Nuss, Johann Cohen Tanugi, John Soo)

@@

\subsection{Training-based Codes}
\label{sec:trainingcodes}
The training-based codes use a variety of algorithms in order to estimate $p(z)$. 
Some aspects of data treatment were left to the individual code runners, for example, whether/how to split the available data with known redshifts into separate training and validation sets.  The treatment of objects not detected in one or more bands was also left to the code runners.  There are varying conventions among training-based codes, and no one prescription dominates in the photo-$z$ literature.  This and other details of the code implementations are described below.

@@

\textsc{ANNz2}\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} is a software package that has the ability to employ several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN). Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can run multiple machine learning algorithms for a single training and outputs photo-$z$'s based on a weighted average of their performances.

\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.  It can also perform classifications, and supports reweighting between samples. The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run. \textsc{ANNz2} presents its photo-$z$ uncertainty different from many codes by using the KNN method: it estimates the photo-$z$ bias between each object and a fixed number of nearest neighbours in parameter space, it then takes the $68$th percentile width of the distribution of the bias. This is based on the implication that objects with similar photometric properties would have similar uncertainties, and therefore the photometric errors of the inputs are not propagated into the code.

@@

\subsubsection{Color-Matched Nearest-Neighbours}
\label{sec:cmnn}
%(Melissa Graham)

The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \citet[][herafter G18]{Graham:17}. This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage. For each galaxy in the test set we identify a color-matched subset of training galaxies, choose one (e.g. the nearest-neighbour or a random selection), and use its known redshift as the photo-$z$. This color-matched subset is identified by first calculating the Mahalanobis distance $D_M$ in color-space between the test galaxy and all training-set galaxies: the difference between the test and a training set galaxy's color divided by the photometric error, summed over all colors (i.e., $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$). Then, the threshold value for $D_M$ that define a good color match is set by the percent point function (PPF): for example, for $N_{\rm dof}=5$, PPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$, the number of degrees of freedom, is the number of colors). For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this color-matched subset of training galaxies, and the standard deviation of the color-matched subset is used as the photo-$z$ uncertainty.

We have applied the nearest-neighbours color-matching photometric redshift estimator described in G18 to the simulated data. Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue. First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment. Second, as in G18 we do apply an initial cut in color to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colors and magnitudes. Third, we set different parameters for the identification of the color-matched subset of training galaxies and the selection of a photometric redshift estimate. In G18 we used a percent point function (PPF) value of 0.68 to identify the color-matched subset of training galaxies and used the redshift of nearest neighbour in color-space as the photo-$z$ estimate. These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return a robust $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in color-space. Since a robust estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator. We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of color-matched training galaxies, but weight them by the inverse of their Mahalanobis distance. This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in color-space. We then use the weights to create the $p(z)$ instead of having the redshift of each color-matched training-set galaxy count equally. To obtain a robust estimate of the $p(z)$ for galaxies with a small number of color-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in color-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:

@@

\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it. Overall, these three changes will yield poorer accuracy photo-$z$ compared to those presented in G18, but they will all have significantly more robust estimates of the $p(z)$, particularly for the brightest test galaxies. This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates. With this work we show how the input parameters should be set in order to return robust $p(z)$ estimates in addition to point value estimates.

@@

\textsc{Delight}\footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} infers photo-$z$'s by using a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift. Generally, machine learning methods rely on representative training data with similar band passes, while template based methods rely on a complete library of templates based on physical models constructed. \textsc{Delight} is constructed in attempt to combine the advantages and eliminate the disadvantages of both template-based and machine learning algorithms: it constructs a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model. The advantage of \textsc{Delight} is that it neither needs representative training data in the same photometric bands, nor does it need detailed galaxy SED models to work.

@@

In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained using the provided training sample. Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands. The default settings of \textsc{Delight} were used, with the exception that the PDF bins were set to be linearly-spaced rather than logarithmic. In this study a flat prior is assumed.

@@

In this paper, we use \textsc{xgboost}~\citep{Chen:16} for the regression part as these techniques scale well for massive data; it should however be noted that \textsc{FlexCode-RF} (also on GitHub), based on Random Forests, generally performs better for smaller data sets. As our basis, we choose a standard Fourier basis. There are two tuning parameters in our $p(z)$ estimate: (i) the number of terms, $I$, in the series expansion in Eq.~\ref{eq::series_expansion}, and (ii) an exponent $\alpha$ that we use to sharpen the computed density estimates $\widehat{f}(z|\x)$, according to $\widetilde{f}(z|\x) \propto \widehat{f}(z|\x)^\alpha$. We split the ``train data'' into a training set (85\%) and a validation set (15\%), and choose both $I$ and $\alpha$ in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq. 5 in][]{Izbicki:17} on the  validation set.

Although {\tt FlexCode} offers a {\em lossless compression} of the photo-$z$ estimates (in this study, one can reconstruct $\widetilde{f}(z|\x)$ exactly at any resolution from estimates of the first 35 coefficients, Eq.~\ref{eq:FlexZBoost_coeffs}, for a Fourier basis $\{\phi_i(z)\}_{i}$), we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$ for easy comparison with other algorithms. Using a higher resolution may yield better results (with no added cost in storage).

@@

The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing the missing values using a simple linear model that estimates the missing variables given the observed ones. The log transformation helps to smooth the long tail distribution of the magnitude errors, which is more stable numerically and makes the optimization process unconstrained. The missing values are imputed by computing the mean of the training set $\mu$ and its covariance $\Sigma$, then we use the following equation to estimate the missing values from the observed ones %\[

\begin{equation}
x_{u} = \mu_{u}+\Sigma_{uo}\Sigma_{oo}^{-1}(x_{o}-\mu_{o}),
\end{equation}  %\]
where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$). This is the optimal expected value of the unobserved variables given the observed ones if the distribution is jointly Gaussian, note that if the variables are independent, i.e. $\Sigma_{uo}=0$, this will reduce to a simple average predictor.

We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.  We do note that \textsc{GPz} is trained using a data set that includes galaxies fainter than those in the test set, which was truncated at $i<25.3$.  This mismatch in training vs.~test results in a slightly different distribution in galaxy photometry and redshift distribution between the training and test set, which most likely degraded results compared to a run where the training set more closely matched the test set.

\subsubsection{METAPhoR}
\label{sec:metaphor}
%(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)
%\textcolor{brown}{general review done on this section}
\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$ point estimates and a reliable PDF for machine learning (ML) based techniques. It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA).
%, already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless of specific implementation, while still taking advantage of the overall code interface.

At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors. The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.

The photometry perturbation law is based on the formula $m_{ij} = m_{ij} + \alpha_{i}F_{ij}*u_{\mu=0,\sigma=1}$, where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.

As introduced, the photo-$z$ point estimate prediction engine of \textsc{METAPhoR} is based on the MLPQNA model, whose photo-$z$ regression training error, used by the quasi Newton learning rule, is based on the least square error and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.

As main prerogative, \textsc{METAPhoR} is able to provide a PDF for ML methods by taking into account the photometric errors provided with data, by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random extractions from the KB. The different test sets, used to produce the PDF, are thus obtained by introducing a proper perturbation, parametrized from the photometric error distribution in each band, on the photometric data populating the original test set \citep{Brescia:18}.

For the present work since it was required to produce a redshift (and a PDF) for each object of the test set we decided to apply a hierarchical kNN to fill the missing detection, it goes without saying that for such points the reliability of PDFs and point estimation is lower. No cross validation has been used.

@@

%In the present work, we use \textsc{SkyNet} both as a regressor (for redshift point estimates) and a classifier (for PDF estimates).
The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with $12$ nodes (the $6$ magnitudes and their errors).
%For the regressor, the output layer is a single linear node, while each hidden layer has $10$ nodes with a tanh activation function.
The classifier is laid out such that the hidden layers have 20:40:40 nodes each, all rectified linear units, and the output layer has $200$ nodes (corresponding to 200 bins for the PDF) activated with a ``softmax'' function so that they automatically sum to $1$.

@@

In the present work, the LSST magnitudes $u,~g,~r,~i$ and colors $u-g,~g-r,~r-i,~i-z,~z-y$ and their associated errors are used in the process of growing 100 trees with a minimum leaf size of 5 (the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation, so we did not use them when constructing our trees). We partitioned our redshift space into 200 bins and smoothed each individual PDF with a smoothing scale of twice the bin size.
%there was a typo here that had 100 bins, I double checked, it is 200 in the final run.

@@

\begin{equation}
p(z) = \frac{1}{N_{ \mathrm train}}\sum_{\mathrm i=1}^{N_{\mathrm train}}z_{\mathrm train}
\end{equation}
That is, we simply set the redshift PDF of every galaxy equal to the normalized $N(z)$ of the training sample.  As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical.  This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.  If the training set was not representative, this estimator would produce biased results, and any attempts to break up the sample into tomographic bins will fail, as every galaxy has an identical $p(z)$.  We will discuss this method and cautions relative to metrics in Section~\ref{sec:caution}.

>>>>

\section{Methods}
\label{sec:pzcodes}
%(Sam Schmidt, Rongpu Zhou, Ibrahim Almosallam, Eric Nuss, Johann Cohen Tanugi, John Soo)

@@

\subsection{Training-based Codes}
\label{sec:trainingcodes}
The training-based codes use a variety of algorithms in order to estimate $p(z)$, specifics of each implementation are described in the subsections.  Some aspects of data treatment were left to the individual code runners, for example, whether/how to split the available data with known redshifts into separate training and validation sets.  Another key difference is the treatment of non-detections in one or more bands.  Some codes choose to ignore a band, others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.  There are varying conventions among training-based codes for treatment of non-detections, and no one prescription dominates in the photo-$z$ literature.  The specific choices for each code affect the results, and contribute to the implicit prior influencing their output.  However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the u-band, and thus should not dominate the code performance differences.

@@

\textsc{ANNz2}\footnote{\url{https://github.com/IftachSadeh/ANNZ}} \citep{Sadeh:16} is a software package that has the ability to employ several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree (BDT) and k-nearest neighbour (KNN). Using the Toolkit for Multivariate Data Analysis (TMVA) with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, it can either run a single machine learning algorithm and return results, or it can run multiple algorithms simultaneously and output photo-$z$'s as a weighted combination of the different algorithms.  In this study, only ANNs were employed.  The redshift PDFs are produced by running an ensemble set of ANNs, each with different random seeds used in initialization of input parameters for training.  Uncertainties for each method are estimated from a KNN-uncertainty estimator \citep{Oyaizu:08}.  The final PDF can either be the ``best'' of the candidate PDFs, or a weighted average of the PDFs based on their error estimates for each of the ensemble members.
%\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.  It can also perform classifications, and supports reweighting between samples.  The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run.

@@

\subsubsection{Colour-Matched Nearest-Neighbours}
\label{sec:cmnn}
%(Melissa Graham)

The nearest-neighbours colour-matching photometric redshift estimator (\textsc{CMNN}) is presented in \citet[][hereafter G18]{Graham:17}. This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage. For each galaxy in the test set we identify a colour-matched subset of training galaxies.  This subset is identified by first calculating the Mahalanobis distance $D_M$ in colour-space between the test galaxy and all training-set galaxies:
\begin{equation} 
D_{\rm M} = \sum^{N_{\rm colours}}\frac{(c_{\rm train}-c_{\rm test})^2}{(\delta c_{\rm test})^2}
\end{equation}
where c colour, $\delta c_{\rm test}$ is the measurement error on the colour, and $N_{\rm colours}$ is the total number of colors (i.e., in our case $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$). Then, we choose a threshold value for $D_M$ that defines the colour match set based on a set value of the percent point function (PPF): for example, for $N_{\rm dof}=5$, choosing a vPPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$ is the number of degrees of freedom, in this case the number of colours). If a galaxy had a non-detection in a band, that band was dropped and $N_{\rm dof}$ was reduced by one in the colour-matching space.  For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this colour-matched subset of training galaxies.

We have applied the nearest-neighbours colour-matching photometric redshift estimator described in G18 to the simulated data. Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue. First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment. Second, as in G18 we do apply an initial cut in colour to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colours and magnitudes. Third, we set different parameters for the identification of the colour-matched subset of training galaxies and the selection of a photometric redshift estimate. In G18 we used a percent point function (PPF) value of 0.68 to identify the colour-matched subset of training galaxies and used the redshift of nearest neighbour in colour-space as the photo-$z$ estimate. These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return an accurate $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in colour-space. Since an accurate estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator. We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of colour-matched training galaxies, but weight them by the inverse of their Mahalanobis distance. This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in colour-space. We then use the weights to create the $p(z)$ instead of having the redshift of each colour-matched training-set galaxy count equally. To obtain a robust estimate of the $p(z)$ for galaxies with a small number of colour-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in colour-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:

@@

\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it. Overall, these three changes will yield less precise photo-$z$ estimates compared to those presented in G18, but they will all have significantly more accurate estimates of the $p(z)$, particularly for the brightest test galaxies. This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates. With this work we show how the input parameters should be set in order to return accurate $p(z)$ estimates in addition to point value estimates.

@@

\textsc{Delight}\footnote{\url{https://github.com/ixkael/Delight}} \citep{Leistedt:17} infers photo-$z$'s by using a data-driven model of latent SEDs and a physical model of photometric fluxes as a function of redshift. Delight models the underlying latent SEDs as a linear combination of a set of pre-defined template SEDs, plus zero mean Gaussian processes with factorized kernels.  Generally, machine learning methods rely on representative training data with similar band passes, while template based methods rely on a complete library of templates based on physical models constructed. \textsc{Delight} is constructed in attempt to combine the advantages and eliminate the disadvantages of both template-based and machine learning algorithms: it constructs a large collection of latent SED templates (or physical flux-redshift models) from training data, with a template SED library as a guide to the learning of the model. The advantage of \textsc{Delight} is that it neither needs representative training data in the same photometric bands, nor does it need detailed galaxy SED models to work.

@@

In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained using the provided training sample. Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands. The default settings of \textsc{Delight} were used, with the exception that the PDF bins were set to be linearly-spaced rather than logarithmic. In this study a flat prior in magnitude/type is assumed.

@@

In this paper, we use \textsc{xgboost}~\citep{Chen:16} for the regression part; it should however be noted that \textsc{FlexCode-RF} (also on GitHub), based on Random Forests, generally performs better for smaller data sets. As our basis, we choose a standard Fourier basis. There are two tuning parameters in our $p(z)$ estimate: (i) the number of terms, $I$, in the series expansion in Eq.~\ref{eq::series_expansion}, and (ii) an exponent $\alpha$ that we use to sharpen the computed density estimates $\widehat{f}(z|\x)$, according to $\widetilde{f}(z|\x) \propto \widehat{f}(z|\x)^\alpha$.  We reserve 15\% of the training set data as a validation set, and choose both $I$ and $\alpha$ in an automated way by minimizing the weighted $L_2$-loss function \citep[Eq. 5 in][]{Izbicki:17} on the  validation set.  While the native storage format for \textsc{FlexCode} encodes the PDF using the coefficients shown in Equation~\ref{eq:FlexZBoost_coeffs}, to match the output format requested of other codes we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$.

%Although {\tt FlexCode} offers a lossless compression of the photo-$z$ estimates (in this study, one can reconstruct $\widetilde{f}(z|\x)$ exactly at any resolution from estimates of the first 35 coefficients, Eq.~\ref{eq:FlexZBoost_coeffs}, for a Fourier basis $\{\phi_i(z)\}_{i}$), we discretize our final estimates into 200 bins linearly spaced in $0 < z < 2$ for easy comparison with other algorithms. Using a higher resolution may yield better results (with no added cost in storage).

@@

The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing any missing magnitude values using a simple linear model that estimates the missing magnitudes given the observed ones. The log transformation helps to smooth the long tail distribution of the magnitude errors, which is more stable numerically and makes the optimization process unconstrained. The missing values are imputed by computing the mean of the training set $\mu$ and its covariance $\Sigma$, then we use the following equation to estimate the missing values from the observed ones %\[

\begin{equation}
x_{u} = \mu_{u}+\Sigma_{uo}\Sigma_{oo}^{-1}(x_{o}-\mu_{o}),
\end{equation}  %\]
where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$). This is the optimal expected value of the unobserved variables given the observed ones if the distribution is jointly Gaussian, note that if the variables are independent, i.e. $\Sigma_{uo}=0$, this will reduce to a simple average predictor.  We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.

\subsubsection{METAPhoR}
\label{sec:metaphor}
%(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)
%\textcolor{brown}{general review done on this section}
\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$ point estimates and a reliable PDF for machine learning (ML) based techniques. It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA).
%, already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.

\textsc{METAPhoR} includes data modules for pre-processing, photo-$z$ estimation, and PDF estimation, and post-processing.  The pre-processing includes a model for perturbation of the photometry that is employed in calculating the PDF of the photo-$z$ estimation errors.  The photometric perturbation is defined as: $m_{ij} = m_{ij} + \alpha_{i}F_{ij}*u_{\mu=0,\sigma=1}$, where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.

%At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors. The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.

%As introduced, the photo-$z$ point estimate prediction engine of \textsc{METAPhoR} is based on the MLPQNA model, whose photo-$z$ regression training error, used by the quasi Newton learning rule, is based on the least square error and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.

As main prerogative, \textsc{METAPhoR} is able to provide a PDF for ML methods by taking into account the photometric errors provided with data, by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random extractions from the KB. The different test sets, used to produce the PDF, are thus obtained by introducing a proper perturbation, parametrized from the photometric error distribution in each band, on the photometric data populating the original test set \citep{Brescia:18}.  For the present work since it was required to produce a redshift (and a PDF) for each object of the test set we decided to apply a hierarchical kNN to replace the missing detections with values based on their neighbors.  The reliability of PDFs and point estimation is lower. No cross validation has been used.

@@

%In the present work, we use \textsc{SkyNet} both as a regressor (for redshift point estimates) and a classifier (for PDF estimates).
The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with $12$ nodes (the $6$ magnitudes and their errors).
%For the regressor, the output layer is a single linear node, while each hidden layer has $10$ nodes with a tanh activation function.
The classifier is laid out such that the hidden layers have 20:40:40 nodes each, all rectified linear units, and the output layer has $200$ nodes (corresponding to 200 bins for the PDF) activated with a ``softmax'' function so that they automatically sum to $1$.  While previous implementations of the code, such as \citet{Sanchez:14} and \citet{Bonnett:15} (see Appendix C.3), implement a ``sliding bin'' smoothing, no such procedure was used in this study.

@@

In the present work, the LSST magnitudes $u,~g,~r,~i$ and colours $u-g,~g-r,~r-i,~i-z,~z-y$ and their associated errors are used in the process of growing 100 trees with a minimum leaf size of 5 (the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation, so we did not use them when constructing our trees). We partitioned our redshift space into 200 bins and smoothed each individual PDF with a smoothing scale of twice the bin size.
%there was a typo here that had 100 bins, I double checked, it is 200 in the final run.

@@

\begin{equation}
p(z) = \frac{1}{N_{ \mathrm train}}\sum_{\mathrm i=1}^{N_{\mathrm train}}z_{\mathrm train}
\end{equation}
That is, we simply set the redshift PDF of every galaxy equal to the normalized $N(z)$ of the training sample.  This estimator is essentially a k nearest-neighbour estimator with k equal to the number of galaxies in the training sample.  As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical, modulo fluctuations due to finite sample size.  This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.  If the training set was not representative, this estimator would produce biased results, and any attempts to break up the sample into tomographic bins will fail, as every galaxy has an identical $p(z)$.  We will discuss this method and cautions relative to metrics in Section~\ref{sec:caution}.
